Towards a Pandemic-Scale Ancestral Recombination Graph of SARS-CoV-2

### 2023-04-21 paper.tex
Update initial Results section

\section{Results}
\subsection{Inferred ARGs} 
To illustrate the use of sc2ts, we employed two different sampling strategies. First, we inferred a densely sampled but time-limited ``wide'' ARG by attaching all the sample genomes with an MSD of 30 days, one daily batch at a time. The inference procedure ran until the genomes sampled on June 30, 2021 were attached to the ARG, resulting in a total of 1.27 million samples. Second, we inferred a ``long'' ARG over a more extended timespan, by attaching at most 1,000 genomes, which were randomly subsampled from each daily batch with an MSD within 30 days. For this ARG, the inference procedure ran until the genomes sampled on June 30, 2022 were attached to the ARG. The long ARG covers an extra year of samples but is more sparsely sampled. June 30, 2022 is the collection date of the final batch of sequences which were preprocessed for the development of sc2ts. Once sc2ts is further developed and tested, we will attach samples collected after that date to infer a fully sampled pandemic-scale ARG.

@@

Most of the nodes in the ARGs represent sample genomes (Table 1). Only a small fraction of the nodes represent recombination events (0.28\% and 0.27\% in the wide ARG and long ARG, respectively), over half of which have no descendant samples (besides the recombinant sequence that triggered insertion of the recombination node) (Table 1). There are over one million mutations on the ARGs, the majority of which occur on a terminal edge (Table 1). Notably, the massive amount of genomic variation data in the wide ARG (1.27 million genomes) and long ARG (657 thousand genomes) are compactly stored in ~57 Mb and ~36 Mb, respectively (less than ten bytes per sample). These ARGs can be loaded for analysis in a Jupyter notebook on a typical commercial laptop in merely a fraction of a second using the Python package tskit (Table 1).

>>>>

\section{Results}
\subsection{Inferred ARGs}
For this prelimary study our goals are to illustrate the utility of
\texttt{sc2ts}, but to also closely study the properties of the
inferred ARGs to inform subsequent development. Thus, rather
than attempting to be comprehensive we chose to work with a
representative subset of the available data, and limited to
inferences that could be performed on a single server in
a few weeks (see below for further details on timings and
computer hardware used). The cut-off dates for sampling were chosen
arbitrarily. We anticipate that further development of the
inference methods and scaling out across multiple servers will
enable inferences at the full pandemic scale using \texttt{sc2ts}
(see Discussion).

To investigate the properties of the inferred ARGs under different
sampling conditions, we inferred two ARGs, which we refer to as the
``Wide'' and ``Long'' ARGs throughout.
The Wide ARG is densely sampled but time-limited and
includes all 1.27 million sample genomes  with a
maximum sampling delay of 30 days (see Methods) up to June 30, 2021.
The Long ARG is more sparsely sampled but
includes data for an additional year (June 30, 2022).
For the long ARG, we use at most 1,000 randomly subsampled
genomes per day from those meeting the maximum sampling delay criterion.

@@

The properties of the inferred ARGs are summarised in Table~\ref{tab:args}.
% TODO more discussion here, making sure we explain anything that might not be
% obvious.
Most of the nodes in the ARGs represent sample genomes
Only a small fraction of the nodes represent recombination events (0.28\% and
0.27\% in the wide ARG and long ARG, respectively), over half of which have no
descendant samples (besides the recombinant sequence that triggered insertion
of the recombination node). There are over one million mutations on
the ARGs, the majority of which occur on a terminal edge. Notably,
the massive amount of genomic variation data in the wide ARG (1.27 million
genomes) and long ARG (657 thousand genomes) are compactly stored in ~57 Mb and
~36 Mb, respectively (less than ten bytes per sample). These ARGs can be loaded
for analysis in a Jupyter notebook on a typical laptop in a
fraction of a second using the Python package tskit.
